---
title: "startup investments"
output:
  github_document: 
    toc: true
date: "2025-05-18"
---

# Preamble/setup

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(showtext)
library(viridis)
library(rnaturalearthdata)
library(sf)
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
     paste("Time for the chunk", options$label, "to run:", res)
    }
  }
}))
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 7)

font_add_google("Lato")
showtext_auto()

custom_theme <- function(...){
  theme_bw() + 
    theme(
      text = element_text(family = "Lato", size = 14),
      title = element_text(face = "bold"),
      strip.text = element_text(face = "bold"),
      plot.caption = element_text(face = "plain"),
      ...
    )
}
theme_set(custom_theme())

options(
  ggplot2.discrete.fill = function(...) scale_fill_brewer(palette = "Set1", ...),
  ggplot2.discrete.colour = function(...) scale_colour_brewer(palette = "Set1", ...)
)

format_mil <- \(x){format(round(x/1e6, 2), trim=TRUE, nsmall=1, big.mark = ",", scientific = FALSE, drop0trailing=T, zero.print=F)}
print_num <- \(x){format(round(x, 2), big.mark = ",",drop0trailing = T, zero.print=F)}
```

```{r import}
# lazy=TRUE here makes readr uses the `vroom` style of processing only the parts of 
# datasets accessed by later code, which can offer significant performance enhancements
options(readr.read_lazy = TRUE)
investors <- read_csv("data/investors.csv")
investments <- read_csv("data/investments.csv")
acquisitions <- read_csv("data/acquisitions.csv")
ipos <- read_csv("data/ipos.csv")
funding_rounds <- read_csv("data/funding_rounds.csv")
funding_round_participants <- funding_rounds |> 
  select(uuid, investor_count) |> 
  left_join(select(investments, funding_round_uuid, investor_uuid, investor_name),
            by = c("uuid" = "funding_round_uuid")) |> 
  group_by(uuid) |> 
  summarize(investor_uuids = list(investor_uuid),
            investor_names = list(investor_name))
funding_rounds <- left_join(
  funding_rounds, funding_round_participants, 
  by = "uuid")
orgs <- read_csv("data/organizations.csv")

# Eventually this should be made into a static/hard-coded table
# But my guess is that false positives like "Banca da Amazonia" don't really
# amount to anything, so it can be a thing done later 
clouds <- investors |> 
  mutate(cloud_giant = case_when(
    str_detect(name, "Google|Alphabet") ~ "Google",
    str_detect(name, "Microsoft") | name == "Azure" ~ "Microsoft", # lots of random "azure" ones, should dig in later to find out which are actually msft
    str_detect(name, "Amazon|AWS") ~ "Amazon",
    T ~ NA_character_
  )) |> 
  drop_na(cloud_giant)

cloud_investments <- investments |> 
  inner_join(
    select(clouds, uuid, cloud_giant),
    by = c("investor_uuid" = "uuid")
    ) |> 
  left_join(
    select(funding_rounds, uuid, raised_amount_usd, investor_count, investment_type, announced_on, org_uuid),
    by = c("funding_round_uuid" = "uuid")) |> 
  mutate(is_lead_investor = case_when(
    !is.na(is_lead_investor) ~ is_lead_investor,
    investor_count == 1 ~ TRUE,
    T ~ NA
  ))

cloud_orgs <- cloud_investments |> 
  select(cloud_giant, investor_name, org_uuid) |> 
  distinct() |> 
  left_join(orgs, by = c("org_uuid" = "uuid"))


active_vcs <- investors |> 
  filter(str_detect(investor_types, "venture_capital")) |> 
  slice_max(order_by = investment_count, n = 10) |>  # other options: ... rank? total_funding_usd is blank
  mutate(criteria = "Very active VC")

active_accels <- investors |> 
  filter(str_detect(investor_types, "accelerator"),
         country_code == "USA") |> 
  slice_max(order_by = investment_count, n = 10) |> 
  mutate(criteria = "Active Accelerator")

# From pitchbook, largest us-based vc firms by aum
big_vc_names <- c(
  "Insight Partners",
  "Sequoia Capital",
  "Andreessen Horowitz",
  "Tiger Global Management",
  "General Catalyst",
  "Accel", 
  "New Enterprise Associates",
  "Lightspeed Venture Partners",
  "Thrive Capital",
  "Greenspring Associates"
)

big_vcs <- investors |> 
  filter(name %in% big_vc_names) |> 
  mutate(criteria = "Big VC")

# or manga or whatever the fuck
# Should refine to get sub-companies
mag_seven <- investors |> 
  filter(name %in% c("Meta", "Apple", "NVIDIA", "Tesla")) |> 
  mutate(criteria = "Other Magnificent Seven")

controls <- bind_rows(
  active_accels,
  big_vcs, 
  mag_seven, 
) |> 
  select(name, criteria, investment_count, uuid)

# large_corps <- orgs |> 
#   filter(primary_role == "company") |> 
#   select(cb_url, category_groups_list) |> 
#   inner_join(investors, by = "cb_url") |> 
#   filter(str_detect(roles, "company"), is.na(investor_types))

control_investments <- investments |> 
  inner_join(select(controls, "uuid", criteria), by = c("investor_uuid" = "uuid")) |> 
  left_join(select(funding_rounds, uuid, raised_amount_usd, investor_count, investment_type, org_uuid, announced_on),
                by = c("funding_round_uuid" = "uuid"))

comparisons <- investors |> 
  select(uuid) |> 
  mutate(criteria = "All investors") |> 
  bind_rows(
    select(clouds, uuid, criteria = cloud_giant),
    select(controls, uuid, criteria)
  ) |> 
  mutate(criteria = fct_relevel(criteria, "Google", "Microsoft", "Amazon", "Other Magnificent Seven", "Big VC", "Active Accelerator", "All investors"),
         # Useful to treat all google investments, etc as coming from one entity
         corp_id = case_when(
           criteria %in% c("Google", "Amazon", "Microsoft") ~ criteria,
           T ~ uuid
         )) |> 
  rename(investor_uuid = uuid)

```

# Figure 1: Big Cloud invests a lot compared to other important investors

```{r fig_1}

comparisons |> 
  left_join(investments, by = "investor_uuid") |> 
  group_by(criteria) |> 
  summarize(n_investments_per_firm = n()/length(unique(corp_id))) |> 
  ggplot(aes(x = fct_relabel(criteria, ~str_wrap(., 10)), y = n_investments_per_firm)) + 
  geom_bar(stat = "identity", fill = RColorBrewer::brewer.pal(3, "Set1")[2]) +
  labs(
    x = NULL, y = "Number of investments per firm",
    caption = "`Per firm` is the number of investments in a given group divided by the number of corporate\n entities in that group. Google, Amazon, and Microsoft have all of their corporate aliases\ntreated as the same group, but the same was not performed for Meta, Netflix, Nvidia, etc." |> str_wrap(80),
    title = "Big Cloud invests way more than other important investors"
  ) 

```

# Figure 1.5: Big Cloud invests more in <x stage> more than others

```{r fig_1b}
types_levels <- c(
  "grant",
  "product_crowdfunding",
  "equity_crowdfunding",
  "pre_seed",
  "seed",
  "non_equity_assistance", 
  "angel", 
  "convertible_note",
  "series_a",
  "series_b",
  "series_c",
  "series_d",
  "series_e",
  "series_f",
  "series_g",
  "series_h",
  "series_i",
  "series_j",
  "initial_coin_offering",
  "post_ipo_equity",
  "post_ipo_secondary",
  "post_ipo_debt",
  "corporate_round",
  "private_equity",
  "debt_financing", 
  "series_unknown",
  "undisclosed",
  "secondary_market"
)

stage_list <- list(
  "Seed, pre-seed, and grants" = c("seed", "pre_seed", "grant", "angel", "non_equity_assistance"),
  "Early-stage" = c("series_a", "series_b", "convertible_note"),
  "Later-stage" = c("series_c", "series_d", "series_e", "series_f", "series_g", "series_h", "series_i", "series_j"),
  "Post-IPO/ICO and corporate rounds" = c("post_ipo_equity", "post_ipo_secondary", "post_ipo_debt", "corporate_round", "private_equity", "non_equity_assistance", "debt_financing", "initial_coin_offering"),
  "Other" = c(
    "undisclosed", "series_unknown", "secondary_market", ""
  )
) 
stages <- tibble(stage_list) |> 
  mutate(stages = fct(names(stage_list), levels = names(stage_list)) |> 
           fct_relabel(~str_wrap(., 10))) |> 
  rename(investment_type = stage_list) |> 
  unnest(investment_type)

comparisons |> 
  left_join(investments, by = "investor_uuid") |> 
  left_join(select(funding_rounds, uuid, investment_type), by = c("funding_round_uuid" = "uuid")) |> 
  mutate(investment_type = fct(investment_type, levels = types_levels)) |> 
  left_join(stages, by = "investment_type") |> 
  drop_na(stages) |> 
  group_by(criteria, stages) |> 
  summarize(n_investments_per_firm = n()/length(unique(corp_id))) |> 
  ggplot(aes(x = fct_relabel(criteria, ~str_wrap(., 10)), fill = stages, y = n_investments_per_firm)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  labs(
    x = NULL, y = "Number of investments per firm",
    fill = NULL,
    caption = "`Per firm` is the number of investments in a given group divided by the number of corporate\n entities in that group. Google, Amazon, and Microsoft have all of their corporate aliases\ntreated as the same group, but the same was not performed for Meta, Netflix, Nvidia, etc." |> str_wrap(80),
    title = "Big Cloud invests at all stages of the startup lifecycle"
  ) + 
  theme(legend.key.spacing.y = unit(0.3, "cm"))
```

# Figure 2: Big Cloud is leading more deals or active as the sole investor more often

```{r fig_2}
comparisons |> 
  left_join(investments, by = "investor_uuid") |> 
  left_join(select(funding_rounds, uuid, org_uuid, investment_type, investor_uuids),
            by = c("funding_round_uuid" = "uuid")) |> 
  mutate(is_sole_investor = map_dbl(investor_uuids, ~length(.) == 1),
         is_lead_investor = is_lead_investor | is_sole_investor) |>
  # Have to do these merges up here because the next summarization happens by umbrella corp instead of a specific corp (like all of google as opposed to google startups or whatever)
  left_join(select(acquisitions, acquirer_uuid, acquiree_uuid) |> 
              distinct() |> 
              mutate(was_acquired = TRUE),
            by = c("investor_uuid" = "acquirer_uuid", "org_uuid" = "acquiree_uuid")) |> 
  left_join(select(ipos, "org_uuid") |> 
              # some have IPOs more than once, like on different stock exchanges or as a result of acquisitions
              distinct() |> 
              mutate(had_ipo = TRUE),
            by = "org_uuid") |> 
  # Ugh have to summarize thtwo  times -- once to our corp IDs, once to big criteria groups
  # to accurately track company-level metrics like repeat investments, acquisitions, IPOs
  group_by(criteria, corp_id, org_uuid) |> 
  summarize(
    is_sole_investor = sum(is_sole_investor, na.rm=T),
    is_lead_investor = sum(is_lead_investor, na.rm=T),
    repeat_investments = n() > 1,
    was_acquired = any(was_acquired),
    had_ipo = any(had_ipo)
  ) |> 
  group_by(criteria) |> 
  summarize(
    "% sole investor" = sum(is_sole_investor, na.rm=T)/n(),
    "% lead investor" = sum(is_lead_investor, na.rm=T)/n(),
    "% repeat investment" = sum(repeat_investments, na.rm=T)/n(),
    # "% acquired" = sum(was_acquired, na.rm=T)/n()
    "% IPO" = sum(had_ipo, na.rm=T)/n()
  ) |> 
  pivot_longer(cols = contains("%")) |> 
  ggplot(aes(x = fct_relabel(criteria, ~str_wrap(., 10)), fill = name, y = value)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  labs(
    x = NULL, y = NULL,
    fill = NULL,
    caption = "Sole investor or lead investor statuses are taken per investment, while repeat investments, acquisitions, or IPOs are assessed for individual companies." |> str_wrap(80)
  ) + 
  theme(legend.key.spacing.y = unit(0.3, "cm"))
```

# Figure 2.5: Co-invested companies are more often in AI

```{r top_sectors_overall}
top_sectors <- cloud_orgs |> 
  select(category_list) |> 
  mutate(category_list = str_split(category_list, ",")) |> 
  unnest(category_list) |> 
  drop_na(category_list) |> 
  count(category_list) |> 
  filter(!(category_list %in% c("Software", "Information Technology"))) |> 
  slice_max(order_by = n, n = 10)
```


```{r coinvestor_heat_map, fig.height = 10}

coinvestors <- cloud_investments |> 
  select(cloud_giant, funding_round_uuid) |> 
  left_join(funding_round_participants, by = c("funding_round_uuid" = "uuid")) |> 
  unnest(cols = investor_uuids) |> 
  group_by(cloud_giant, investor_uuids) |> 
  count(name = "n_deals") |> 
  group_by(cloud_giant) |> 
  mutate(pct_deals = round(100 * n_deals / n(), 2)) |> 
  left_join(select(investors, uuid, name), by = c("investor_uuids" = "uuid")) |> 
  # anti-join to filter out 
  anti_join(select(clouds, cloud_giant, name),
            by = c("cloud_giant", "name")) |> 
  arrange(desc(pct_deals)) |> 
  slice_max(order_by = pct_deals, n = 20, with_ties = FALSE) |> 
  mutate(deals = paste0(n_deals, " (", pct_deals, "%)"))

ggplot(coinvestors, aes(x = cloud_giant, y = name, fill = pct_deals)) + 
  geom_tile() + 
  scale_fill_viridis(option = "B") + 
  labs(y = NULL, x = NULL, fill = "% of deals made\nwith coinvestor", 
       title = "Heatmap of co-investment by Big Three",
       subtitle = "It looks splotchy because of only a moderate amount of overlap")

```

# Figure 3: The Big Cloud investors find their own niches


```{r big_cloud_sectors, fig.width = 10}
big_cloud_top_sector <- cloud_investments |> 
  left_join(select(cloud_orgs, org_uuid, category_list) |> distinct(), by = c("org_uuid" = "org_uuid")) |> 
  mutate(category_list = str_split(category_list, ",")) |> 
  unnest(category_list) |> 
  filter(category_list %in% top_sectors$category_list) |> 
  # mutate(date = floor_date(announced_on, "3 months")) |> 
  group_by(cloud_giant, category_list) |> 
  summarize(n_deals = n(),
            amt = sum(raised_amount_usd, na.rm=T)) |> 
  mutate(frac_deals = n_deals / sum(n_deals))


ggplot(big_cloud_top_sector, aes(x = str_wrap(category_list, 10), y = n_deals,
                                 fill = cloud_giant)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(
    x = NULL, y = "Number of deals", fill = NULL
  ) + 
  theme(legend.position = "top")
```

```{r big_cloud_sectors_prop, fig.width = 15}
ggplot(big_cloud_top_sector, aes(x = str_wrap(category_list, 10), y = frac_deals,
                                 fill = cloud_giant)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(
    x = NULL, y = "Fraction of total deals made by corporation", fill = NULL
  ) + 
  theme(legend.position = "top")
```

# Figure 3.2: Gov tech

```{r govtech}
govtech_list <- list(
  "Government and Public Sector" = c(
    "Government",
    "GovTech",
    "CivicTech"
  ),
  "Public Safety and Law Enforcement" = c(
    "Public Safety",
    "Law Enforcement",
    "Corrections Facilities"
  ),
  "Military" = c(
    "National Security",
    "Homeland Security",
    "Military"
  ),
  "Municipal and Public Works" = c(
    "Waste Management",
    "Pollution Control",
    "Water",
    "Recycling",
    "Power Grid"
  ),
  "Transport and infra" = c(
    "Public Transportation",
    "Water Transportation",
    "Railroad",
    "Ports and Harbors"
  ),
  "Public Health" = c(
    "Addiction Treatment",
    "Homeless Shelter",
    "Social Assistance"
  )
  # "Education" = c(
  #   "Education", 
  #   "Higher Education",
  #   "Universities",
  #   "Primary Education",
  #   "Secondary Education",
  #   "Charter Schools",
  #   "Special Education",
  #   "STEM Education"
  # )
)

govtech <- tibble(govtech_list) |> 
  mutate(govtech_categories = names(govtech_list)) |> 
  unnest(govtech_list) |> 
  rename(category_list = govtech_list)

govtech_investments <- comparisons |> 
  left_join(investments, by = "investor_uuid") |> 
  group_by(criteria) |> 
  mutate(total_criteria_deals = n()) |> 
  ungroup() |> 
  left_join(select(funding_rounds, uuid, org_uuid), by = c("funding_round_uuid" = "uuid")) |> 
  left_join(select(orgs, uuid, category_list) |> distinct(), by = c("org_uuid" = "uuid")) |> 
  mutate(category_list = str_split(category_list, ",")) |> 
  unnest(category_list) |> 
  right_join(govtech, by = "category_list") |> 
  #make sure counting each funding round once per govtech category
  select(criteria, total_criteria_deals, govtech_categories, funding_round_uuid) |> 
  distinct() |> 
  drop_na(criteria) |> 
  left_join(select(funding_rounds, uuid, raised_amount_usd), by = c("funding_round_uuid" = "uuid")) |> 
  group_by(criteria, total_criteria_deals, govtech_categories)|> 
  summarize(n_deals = n(),
            amt = sum(raised_amount_usd, na.rm=T)) |> 
  mutate(frac_deals = 100 * n_deals / total_criteria_deals)

ggplot(govtech_investments,
       aes(x = str_wrap(govtech_categories, 10),
           y = frac_deals,
           fill = criteria)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(
    x = NULL, y = "% of total deals made by investor or investor group", fill = NULL
  ) + 
  theme(legend.position = "top")
```

# Figure 4: Big Cloud and other major investors have taken a strong pivot to AI

By # of deals:

```{r cloud_ai_deals}
cloud_ai <- cloud_orgs |> 
  mutate(is_ai = str_detect(category_list, "Artificial Intelligence")) |> 
  select(org_uuid, is_ai) |> 
  distinct()
  
cloud_investments |> 
  left_join(cloud_ai, by = "org_uuid") |> 
  drop_na(is_ai) |> 
  mutate(date = floor_date(announced_on, "3 months")) |> 
         # category = paste0(cloud_giant, ifelse(is_ai, " - Not AI", " - AI"))) |> 
  group_by(date, is_ai, cloud_giant) |> 
  count() |> 
  ggplot(aes(x = date, y = n, linetype = is_ai, color = cloud_giant)) + 
  geom_line() +
  scale_x_date(limits = c(as.Date("2014-01-01"), as.Date("2026-01-01")))

```


```{r cloud_ai_amt}
  
cloud_investments |> 
  left_join(cloud_ai, by = "org_uuid") |> 
  drop_na(is_ai) |> 
  mutate(date = floor_date(announced_on, "3 months")) |> 
         # category = paste0(cloud_giant, ifelse(is_ai, " - Not AI", " - AI"))) |> 
  group_by(date, is_ai, cloud_giant) |> 
  summarize(total_amt = sum(raised_amount_usd, na.rm=T)) |> 
  ggplot(aes(x = date, y = total_amt, linetype = is_ai, color = cloud_giant)) + 
  geom_line() + 
  scale_y_continuous(labels = \(x){paste0("$", format(x/1e6, nsmall=1, big.mark = ",", scientific = FALSE))}) + 
  scale_x_date(limits = c(as.Date("2014-01-01"), as.Date("2026-01-01")))

```


```{r amt_cloud}
cloud_investments |> 
  left_join(cloud_ai, by = "org_uuid") |> 
  drop_na(is_ai) |> 
  mutate(date = floor_date(announced_on, "3 months")) |> 
         # category = paste0(cloud_giant, ifelse(is_ai, " - Not AI", " - AI"))) |> 
  group_by(date, is_ai, cloud_giant) |> 
  summarize(total_amt = sum(raised_amount_usd, na.rm=T)) |> 
  ggplot(aes(x = date, y = total_amt, linetype = is_ai, color = cloud_giant)) + 
  geom_line() + 
  scale_y_continuous(labels = \(x){paste0("$", format(x/1e6, nsmall=1, big.mark = ",", scientific = FALSE))}, limits = c(0, 100000000000)) + 
  scale_x_date(limits = c(as.Date("2014-01-01"), as.Date("2026-01-01")))

```


# Figure 5: Big Cloud is investing in all parts of the supply chain

```{r supply_chain_cleaning}
supply_chain <- lst(
  "Hardware" = c(
    "Application Specific Integrated Circuit (ASIC)",
    "GPU",
    "Field-Programmable Gate Array (FPGA)"
  ),
  "Internet Infra" = c(
    "Wired Telecommunications",
    "Network Hardware",
    "Communication Hardware",
    "ISP",
    "Content Delivery Network",
    "Satellite Communication",
    "IT Infrastructure"
  ),
  # Research = c(
  #   "Machine Learning",
  #   "Artificial Intelligence (AI)",
  #   "Generative AI",
  #   "Natural Language Processing",
  #   "Computer Vision"
  # ),
  Data = c(
    "Big Data",
    "Data Mining",
    "Database",
    "Data Integration",
    "Geospatial"
  ),
  # "Responsible AI" = c(
  #   "Compliance",
  #   "Data Governance"
  # ),
  Cloud = c(
    "Cloud Storage",
    "Cloud Infrastructure",
    "Cloud Computing",
    "Data Management",
    "Private Cloud"
  ),
  "As a Service" = c(
    "SaaS", "IaaS", "PaaS", "Developer APIs"
  ),
  "Developer Services" = c(
    "Network Security",
    "Cloud Security",
    "Network Security",
    "DevOps",
    "Developer Tools",
    "Developer Platform"
    ),
  #"Business-facing" = c("B2B"),
  # "Consumer-facing" = c(
  #   "Mobile Apps", "Internet of Things", "Consumer Electronics", "Virtual Reality", "Augmented Reality"
  #   )
)

supply_chain <- supply_chain |> 
  tibble() |> 
  mutate(component = names(supply_chain) |> factor(levels = names(supply_chain))) |> 
  unnest(supply_chain)

individual_comparisons <- bind_rows(
  mutate(clouds, name = cloud_giant),
  mag_seven, 
  slice_head(big_vcs, n = 2)
) |> 
  select(name, investor_uuid = uuid) |> 
  mutate(name = fct_relevel(name, "Google", "Microsoft", "Amazon", "Apple", "Meta", "NVIDIA", "Accel", "Sequoia Capital"))

supply_chain_investments <- investments |> 
  select(investor_uuid, funding_round_uuid, is_lead_investor) |> 
  left_join(select(funding_rounds, uuid, org_uuid, raised_amount_usd), by = c("funding_round_uuid" = "uuid")) |> 
  
  left_join(select(orgs, uuid, category_list) |> distinct(), by = c("org_uuid" = "uuid")) |> 
  right_join(individual_comparisons, by = "investor_uuid") |> 
  mutate(category_list = str_split(category_list, ",")) |> 
  unnest(category_list) |> 
  full_join(supply_chain, by = c("category_list" = "supply_chain")) |> 
  select(-category_list) |> 
  distinct() # Since a single company could have two relevant categories from a single supply chain component
```

## 5a: # of deals in a given portion of the supply chain

```{r supply_tile}
supply_chain_investments |> 
  drop_na(component) |> 
  group_by(name, component) |> 
  count() |> 
  ggplot() + 
  geom_tile(aes(x = name, y = component, fill = n)) + 
  scale_fill_viridis()
```

## 5b: # of deals in given supply chain component as proportion of all deals

```{r supply_tile_prop}
supply_chain_investments |>
  group_by(name) |> 
  mutate(denominator = n()) |> 
  drop_na(component) |> 
  group_by(name, component) |> 
  summarize(prop_deals = n()/denominator) |> 
  ggplot() + 
  geom_tile(aes(x = name, y = component, fill = prop_deals)) + 
  scale_fill_viridis()
```

## 5: Total size of all participated funding rounds


```{r size_tile}
supply_chain_investments |>
  filter(raised_amount_usd != 40000000000) |> 
  group_by(name) |> 
  mutate(denominator = sum(raised_amount_usd, na.rm=T)) |> 
  drop_na(component) |> 
  group_by(name, component) |> 
  summarize(prop_deals = sum(raised_amount_usd, na.rm=T)/denominator) |> 
  ggplot() + 
  geom_tile(aes(x = name, y = component, fill = prop_deals)) + 
  scale_fill_viridis() + 
  labs(
    fill = "Prop of raised amount in all\nparticipated rounds",
    caption = "Excludes Microsoft's $40B deal with OpenAI"
  )
```

## 5b: % of lead deals

```{r lead_tile}

supply_chain_lead_sole <- supply_chain_investments |> 
  left_join(funding_round_participants, by = c("funding_round_uuid" = "uuid")) |> 
  mutate(
    is_sole_investor = map_lgl(investor_names, ~length(.) == 1),
    is_lead_investor = is_lead_investor | is_sole_investor
  ) |> 
  drop_na(component) |> 
  group_by(name, component) |> 
  summarize(
    is_sole_investor = sum(is_sole_investor, na.rm=T)/n(),
    is_lead_investor = sum(is_lead_investor, na.rm=T)/n()
    )

supply_chain_lead_sole |> 
  ggplot() + 
  geom_tile(aes(x = name, y = component, fill = 100 * is_lead_investor)) + 
  scale_fill_viridis() + 
  labs(fill = "% of time lead investor", x = NULL, y = NULL,
       title = "big cloud investment leads by sector") + 
  theme(legend.key.height = unit(2, "cm"))
```

## 5c: % of sole investor

```{r sole_tile}
supply_chain_lead_sole |> 
  ggplot() + 
  geom_tile(aes(x = name, y = component, fill = is_sole_investor)) + 
  scale_fill_viridis()
```


## 5d: % of repeat investor

# Figure 6: Big cloud has especially salient interests in X regions

```{r cloud_regions}

data(countries)
country_shapes <- countries50 |>
  filter(region_wb != "Antarctica") |> 
  select(region_wb, country_name = name,iso_a3)

cloud_investments |> 
  left_join(select(funding_rounds, country_code, uuid), by = c("funding_round_uuid" = "uuid")) |> 
  left_join(st_drop_geometry(country_shapes), by = c("country_code" = "iso_a3")) |> 
  group_by(country_code) |> 
  count() |> 
  full_join(select(country_shapes, iso_a3), by = c("country_code" = "iso_a3")) |> 
  mutate(n = case_when(
    country_code == "USA" ~ NaN,
    is.na(n) ~ 0,
    T ~ n)) |> 
  st_as_sf() |> 
  ggplot(aes(fill = n)) + 
  geom_sf() + 
  scale_fill_viridis() +
  labs(title = "Number of deals in a given country")
```

```{r cloud_region_map}
investments |> 
  left_join(select(funding_rounds, country_code, uuid), by = c("funding_round_uuid" = "uuid")) |> 
  left_join(st_drop_geometry(country_shapes), by = c("country_code" = "iso_a3")) |> 
  group_by(country_code) |> 
  count() |> 
  full_join(select(country_shapes, iso_a3), by = c("country_code" = "iso_a3")) |> 
  mutate(n = case_when(
    country_code == "USA" ~ NaN,
    is.na(n) ~ 0,
    T ~ n)) |> 
  st_as_sf() |> 
  ggplot(aes(fill = n)) + 
  geom_sf() + 
  scale_fill_viridis()
```

```{r cloud_region_bar, fig.width = 12}

top_countries <- funding_rounds |> 
  filter(country_code != "USA") |> 
  group_by(country_code) |> 
  summarize(raised_amount_usd = sum(raised_amount_usd, na.rm=T)) |> 
  slice_max(order_by = raised_amount_usd, n = 15) |> 
  select(-raised_amount_usd)

comparisons |> 
  left_join(investments, by = "investor_uuid") |> 
  filter(!(criteria %in% c(
    "Active Accelerator", "Big VC"
  ))) |> 
  group_by(criteria) |> 
  mutate(total_n = n()) |> 
  left_join(select(funding_rounds, country_code, uuid), by = c("funding_round_uuid" = "uuid")) |> 
  left_join(st_drop_geometry(country_shapes), by = c("country_code" = "iso_a3")) |> 
  filter(country_code %in% top_countries$country_code) |> 
  drop_na(country_name) |> 
  group_by(country_name, criteria) |> 
  summarize(prop_n = n() / total_n) |> 
  ggplot(aes(x = fct_reorder(country_name, prop_n) |>
               fct_relabel(~str_wrap(., 10)),
             fill = criteria, y = 100 * prop_n)) + 
  geom_bar(position = "dodge", stat = "identity") + 
  theme(legend.position = "bottom") + 
  labs(
    x = NULL, y = "Percent of deals within group",
    title = "Each member of the Big Cloud has distinct preferences for countries"
  )

```


There seems to be Third World development / investment arms maintained by the big 3 cloud:

- https://www.crunchbase.com/organization/umbono
- Google’s Africa Investment Fund investment in Series C - Moniepoin
- NCAIR-Google AI Fund: Nigeria investments
- Alibaba?

# Figure 7: The Big 3 clouds have different regional preferences (3 maps, one per cloud company)

```{r region_big_cloud_map, fig.width = 10}

country_cloud <- expand_grid(
  iso_a3 = country_shapes$iso_a3,
  cloud_giant = unique(clouds$cloud_giant)
)

cloud_investments |> 
  left_join(select(funding_rounds, country_code, uuid), by = c("funding_round_uuid" = "uuid")) |> 
  left_join(st_drop_geometry(country_shapes), by = c("country_code" = "iso_a3")) |> 
  group_by(cloud_giant) |> 
  mutate(total_n = n()) |> 
  group_by(cloud_giant, total_n, country_code) |> 
  summarize(prop_n = n()) |> 
  mutate(prop_n = prop_n / total_n) |> 
  # To ensure even if a cloud giant made 0 deals to a country it's still mapped
  full_join(country_cloud, by = c("cloud_giant", "country_code" = "iso_a3")) |> 
  full_join(select(country_shapes, iso_a3), by = c("country_code" = "iso_a3")) |> 
  drop_na(cloud_giant) |> 
  mutate(prop_n = case_when(
    country_code == "USA" ~ NaN,
    is.na(prop_n) ~ 0,
    T ~ prop_n)) |> 
  st_as_sf() |> 
  ggplot(aes(fill = prop_n)) + 
  geom_sf() + 
  scale_fill_viridis() + 
  facet_wrap(~cloud_giant, ncol = 1) + 
  labs(
    fill = "Prop. of all deals\nmade by corporation"
  ) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(), 
        axis.line = element_blank())

```

# Figure 8:  Over time, each Big Cloud invests moreso into certain regions (line graph)

Something wrong with this graph -- doesn't make sense to display for the moment. 

```{r region_time, eval = F}
comparisons |> 
  mutate(criteria = case_when(
    criteria %in% c("Google", "Amazon", "Microsoft") ~ "Big Cloud",
    T ~ criteria)) |> 
  full_join(expand_grid(
    criteria = unique(comparisons$criteria),
    year = 2015:2024
  ), by = "criteria") |> 
  filter(!(criteria %in% c("Active Accelerator", "Big VC"))) |> 
  left_join(investments, by = "investor_uuid") |> 
  select(criteria, funding_round_uuid) |> 
  left_join(
    select(funding_rounds, uuid, org_uuid,
           announced_on, country_code, raised_amount_usd),
            by = c("funding_round_uuid" = "uuid")) |> 
  filter(country_code != "USA", year(announced_on) != 2025) |> 
  left_join(country_shapes, by = c("country_code" = "iso_a3")) |> 
  mutate(date = floor_date(announced_on, "1 year")) |> 
  drop_na(region_wb) |> 
  group_by(criteria, region_wb, date) |> 
  summarize(raised_amount_usd = sum(raised_amount_usd, na.rm=T)/n()) |> 
  ggplot(aes(x = date, y = raised_amount_usd, color = region_wb)) + 
  geom_line(size = 1) +
  scale_y_log10(labels = \(x){paste0("$", format_mil(x))}) + 
  scale_x_date(limits = c(as.Date("2010-01-01"), as.Date("2024-01-01")), 
               expand = expansion(add = 0),
               date_breaks = "1 year",
               date_labels = "%Y") + 
  theme(legend.position = "bottom") + 
  labs(
    x = NULL, y = "Total raised amount", color = NULL,
    title = "Investment internationally trends upwards over time"
    ) + 
  facet_wrap(~criteria, ncol = 1)

```


# Table: Big Cloud has especially salient interests in X regions

```{r cloud_region_table}
cloud_orgs |> 
  group_split(cloud_giant) |> 
  map(\(corp_group){
    total_n <- length(unique(corp_group$org_uuid))
    cloud_giant_name <- unique(corp_group$cloud_giant)
    top_ten <- corp_group |> 
      count(country_code) |> 
      slice_max(order_by = n, n = 10)
    all_others <- corp_group |> 
      filter(!(country_code %in% top_ten$country_code)) |> 
      count() |> 
      mutate(country_code = "All others for this cloud corp")
    bind_rows(top_ten, all_others) |> 
      mutate(n = paste0(n, " (", round(100*n/total_n, 2), "%)")) |> 
      rename(!!cloud_giant_name := n)
  }) |> 
  reduce(full_join, by = "country_code") |> 
  kable(caption = "Summary statistics by cloud corp")

```

# Table of investments by top sectors

```{r sector_table}

cloud_sectors <- cloud_investments |> 
  left_join(select(cloud_orgs, org_uuid, category_list) |> distinct(),
            by = c("org_uuid" = "org_uuid")) |> 
  mutate(category_list = str_split(category_list, ",")) |> 
  unnest(category_list) |> 
  filter(category_list %in% top_sectors$category_list) |> 
  group_by(cloud_giant, category_list) |> 
  summarize(n_deals = n(), amt = sum(raised_amount_usd, na.rm=T))
  
  
cloud_sectors |> 
  mutate(stat = paste0(n_deals, " ($", amt |> round(3) |> format_mil(), ")")) |> 
  select(-n_deals, -amt) |> 
  pivot_wider(names_from = cloud_giant, values_from = stat) |> 
  kable()
  
```

# Table: Coinvesting investors

```{r coinvestors_table}


coinvestors |> 
  select(-investor_uuids, -n_deals, -pct_deals) |> 
  pivot_wider(names_from=cloud_giant, values_from=deals) |> 
  mutate(across(where(is.character), ~ifelse(is.na(.), "", .))) |> 
  kable()
```

# Appendix: Aliases of cloud companies

```{r aliases}

cloud_investments |>
  group_by(cloud_giant, investor_name) |> 
  count() |> 
  group_by(cloud_giant) |> 
  arrange(n) |> 
  summarize(aliases = paste0(investor_name, " (", n, ") ", collapse = ", ")) |> 
  rename("Cloud corp" = cloud_giant,
         "Aliases (# deals)" = aliases) |> 
  kable()

```

# Appendix: Important investors to compare against

```{r baseline}
controls |> 
  select(-uuid) |> 
  kable()
```


```{r summary_stats, eval = FALSE, time_it = FALSE}

# convoluted but couldn't think of another way lol 
cloud_investments |> 
  rename(group_name = cloud_giant) |> 
  bind_rows(mutate(investments, group_name = "All investments") |> 
              left_join(
                select(funding_rounds, uuid, raised_amount_usd, investor_count, investment_type, org_uuid, announced_on),
                by = c("funding_round_uuid" = "uuid")),
            control_investments |> mutate(group_name= criteria)
            ) |> 
  group_split(group_name) |> 
  map(\(corp_group){
    
    group_name <- unique(corp_group$group_name)
    repeat_orgs <- corp_group |> 
      select(investor_uuid, org_uuid, funding_round_uuid) |> 
      distinct() |> 
      # Takes a long time for all investments
      group_by(investor_uuid, org_uuid) |> 
      filter(n() > 1) |> 
      pull(org_uuid) |> 
      unique() |> 
      length()
    n_orgs <- length(unique(corp_group$org_uuid))
    org_cats <- corp_group |> 
      select(org_uuid) |> 
      distinct() |> 
      left_join(select(orgs, uuid, category_list), by = c("org_uuid" = "uuid")) |> 
      pull(category_list)
    
    tribble(
      ~Statistic, ~Value,
      "Total number of deals", nrow(corp_group),
      "Total dollar amount of participated deals (in millions, including other investors)", sum(corp_group$raised_amount_usd, na.rm=T)/1e6,
      "Number of investing aliases", length(unique(corp_group$investor_name)),
      # "Most popular stage of investment", sort(table(corp_group$investment_type), descending = T)[1] |>  names(), # shitty way to do that lol
      '% of time that they “lead” funding rounds', 100*sum(corp_group$is_lead_investor, na.rm=T)/length(corp_group$is_lead_investor),
      '% of time that they are the sole investor', 100*sum(corp_group$investor_count == 1, na.rm=T)/length(corp_group$is_lead_investor),
      "% of companies receiving repeat investments", 100*repeat_orgs/length(unique(corp_group$org_uuid)),
      "Median size of participating deal", median(corp_group$raised_amount_usd, na.rm=T)/1e6,
      "Companies receiving investment", n_orgs,
      "% of invested companies in open-source", 100*sum(str_detect(org_cats, "Open Source"), na.rm=T)/n_orgs,
      "% of invested companies in AI", 100*sum(str_detect(org_cats, "Artificial Intelligence"), na.rm=T)/n_orgs,
      "% of invested companies in Crypto", 100*sum(str_detect(org_cats, "Cryptocurrency"), na.rm=T)/n_orgs,
    ) |> 
      mutate(Value = print_num(Value)) |> 
      rename(!!group_name:= Value)
  }) |> 
  reduce(left_join, by = "Statistic") |> 
  kable(caption = "Summary statistics by corp group")

```

# Appendix: Table of investment counts by series and investor group

```{r series_table}

funding_rounds |> 
  rename(funding_round_uuid = uuid) |> 
  left_join(investments, by = "funding_round_uuid") |>
  left_join(comparisons, by = "investor_uuid") |> 
  select(investment_type, investment_uuid = uuid, criteria, raised_amount_usd, corp_id) |> 
  group_by(investment_type, criteria) |> 
  summarize(
    avg_raised = sum(raised_amount_usd, na.rm=T)/length(unique(corp_id)),
    n_deals = round(n()/length(unique(corp_id)), 2)
  ) |> 
  mutate(value = ifelse(is.na(avg_raised), NA_character_, paste0("$", format_mil(avg_raised), " (", n_deals, ")"))) |>
  select(-avg_raised, -n_deals) |> 
  pivot_wider(names_from = criteria, values_from = value) |> 
  left_join(stages, by = "investment_type") |> 
  drop_na(stages) |> 
  mutate(stages = fct_relabel(stages, ~str_replace_all(., "\n", " "))) |> 
  select("Investment Type" = investment_type, Stage = stages, levels(comparisons$criteria)) |> 
  arrange(Stage) |> 
  kable()

  # write_csv("~/Downloads/stages.csv")

```

# Appendix: Largest overall investments by cloud companies

These are outliers, generally -- 

```{r largest_investments}
cloud_investments |> 
  select(name, announced_on, raised_amount_usd) |> 
  slice_max(order_by = raised_amount_usd, n = 20) |> 
  arrange(desc(raised_amount_usd)) |> 
  mutate("Raised amount, millions" = format(round(raised_amount_usd / 1e6, 2), big.mark = ",",drop0trailing = T, zero.print=F)) |> 
  select(-raised_amount_usd) |> 
  kable()
```


# Appendix: When does the Big Cloud invest in the same companies?

```{r big_cloud_coinvestments, eval=F}
received_multiple <- cloud_investments |> 
  select(org_uuid, cloud_giant) |> 
  distinct() |> 
  group_by(org_uuid) |> 
  filter(n() > 1) |> 
  select(-cloud_giant) |> 
  distinct()

received_multiple |> 
  ungroup() |> 
  left_join(cloud_investments, by = "org_uuid") |> 
  select(funding_round_uuid) |> 
  distinct() |> 
  left_join(funding_rounds, by = c("funding_round_uuid" = "uuid")) |> 
  left_join(select(orgs, uuid, category_groups_list, category_list), by = c("org_uuid" = "uuid")) |> 
  mutate(investor_names = map_chr(investor_names, ~paste0(., collapse = ", "))) |> 
  select(org_name, name, announced_on, investor_names, raised_amount_usd, category_groups_list, category_list) |> 
  write_csv("~/Downloads/multiple.csv")
```

<!-- # tables for david to work through -->

```{r all_cloud_investments, eval = FALSE}
cloud_investments_audit <- cloud_investments |> 
  mutate(is_sole_investor = investor_count == 1) |> 
  select(cloud_giant, funding_round_uuid, is_lead_investor, is_sole_investor) |> 
  left_join(select(funding_rounds, uuid, org_uuid), by = c("funding_round_uuid" = "uuid")) |> 
  left_join(select(orgs, name, uuid, category_list), by = c("org_uuid" = "uuid")) |> 
  mutate(category_list = str_split(category_list, ",")) |> 
  unnest(category_list) |> 
  left_join(supply_chain, by = c("category_list" = "supply_chain")) |> 
  group_by(cloud_giant, name, org_uuid, is_lead_investor, is_sole_investor) |> 
  summarize(category_list = paste0(category_list, collapse = ","),
            component = paste0(unique(component[!is.na(component)]), collapse = ",")) |> 
  group_by(cloud_giant, name, org_uuid, category_list, component) |> 
  summarize(times_lead_investor = sum(is_lead_investor, na.rm=T),
            times_sole_investor = sum(is_sole_investor, na.rm=T),
            times_investor = n()
            ) |> 
  ungroup() 
cloud_investments_audit |> 
  write_csv("~/Downloads/cloud_investments_audit.csv")
  
```


```{r non_equity, eval = F}
non_equity <- cloud_investments |> 
  mutate(is_sole_investor = investor_count == 1) |> 
  select(cloud_giant, funding_round_uuid, is_lead_investor, is_sole_investor) |> 
  left_join(select(funding_rounds, uuid, org_uuid, investment_type, announced_on, raised_amount_usd, contains("url")),
            by = c("funding_round_uuid" = "uuid")) |> 
  filter(investment_type == "non_equity_assistance") |> 
  left_join(select(orgs, org_name = name, uuid, category_list), by = c("org_uuid" = "uuid")) 

non_equity |> 
  write_csv("~/Downloads/non_equity.csv")
# name of company, name of cloud, cb url, url, 
```

```{r supply_chain_leads, eval=F}

supply_chain_lead_sole |>
  select(-is_sole_investor) |>
  pivot_wider(values_from=is_lead_investor, names_from=name) |>
  write_csv("~/Downloads/supply_chain_leads.csv")

```
